% Document class and font size
\documentclass[a4paper,9pt]{extarticle}

% Packages
\usepackage[utf8]{inputenc} % For input encoding
\usepackage{geometry} % For page margins
\geometry{a4paper, margin=1in} % Set paper size and margins
\usepackage{titlesec} % For section title formatting
\usepackage{enumitem} % For itemized list formatting
\usepackage{hyperref} % For hyperlinks

% Formatting
\setlist{noitemsep} % Removes item separation
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}[\titlerule] % Section title format
\titlespacing*{\section}{0pt}{\baselineskip}{\baselineskip} % Section title spacing

% Begin document
\begin{document}

% Disable page numbers
\pagestyle{empty}

% Header
\begin{center}
\textbf{\Large James Dolan}\\[3pt] % Name
\textbf{Resume}\\[1pt] % CV
Hauppauge, New York| \href{mailto:james.dolan1@students.mq.edu.au}{james.dolan1@students.mq.edu.au} | (631) 724-2127 | \href{https://ncatlab.org/nlab/show/James+Dolan}{ncatlab.org/nlab/show/James+Dolan} | % Country Citizen % Contact info
\end{center}

%------------------------

% Education Section
\section*{POSITIONS}

\noindent
\newline
1982-1986
SUNY Buffalo mathematics PhD student / teaching assistant 
(studying under William Lawvere and Stephen Schanuel).

\noindent
1993-1996
UC Riverside mathematics PhD student / teaching assistant 
(studying under John Baez).

\noindent
1997
Video co/dec programmer for a startup in Riverside, California.

\noindent
1998
Software engineer for Sanga inc. in Montreal.

\noindent
2001 
Macquarie University mathematics research associate 
(working with Ross Street).

\noindent
1993-2010
Research collaboration with John Baez at UC Riverside.

\noindent
2015-2017
Macquarie University mathematics PhD student (studying under Ross Street).

\noindent
2018
Consultant on patents for data-encryption and user-authentication.


%\noindent
%\textbf{Degree Name} \\
%Year of completion: XXXX \\
%Institution \\
%Department  \\
%Maecenas ultricies augue a turpis viverra iaculis. Fusce a lorem est. Phasellus a metus dignissim, commodo eros eget, tempor ipsum. Pellentesque vel posuere ante, vitae dignissim ligula.

%%------------------------
%
%% Employment Section
%\section*{EMPLOYMENT}
%\noindent
%\newline
%\textbf{Position Title} \\
%Year-Year: XXXX-XXXX \\
%Institution \\
%Department \\
%Maecenas ultricies augue a turpis viverra iaculis. Fusce a lorem est. Phasellus a metus dignissim, commodo eros eget, tempor ipsum. Pellentesque vel posuere ante, vitae dignissim ligula. Maecenas maximus lorem orci, ac semper sapien maximus et. \\
%
%\noindent
%\textbf{Position Title} \\
%Year-Year: XXXX-XXXX \\
%Institution \\
%Department \\
%Maecenas ultricies augue a turpis viverra iaculis. Fusce a lorem est. Phasellus a metus dignissim, commodo eros eget, tempor ipsum. Pellentesque vel posuere ante, vitae dignissim ligula. Maecenas maximus lorem orci, ac semper sapien maximus et.

%------------------------

% Publications Section
\section*{SELECTED PUBLICATIONS}

\noindent
Baez, John C., and James Dolan. 
"Higher‐dimensional Algebra and Topological Quantum Field Theory." 
Journal of Mathematical Physics 36.11 (1995): 6073-6105.

\noindent
Baez, John C., and James Dolan. 
"Higher-Dimensional Algebra III. n-Categories and the Algebra of Opetopes." 
Advances in Mathematics 135.2 (1998): 145-206.

\noindent
Baez, John C., and James Dolan. "Categorification." 
Higher Category Theory, eds. Ezra Getzler and Mikhail
Kapranov, Contemp. Math. 230, American Mathematical Society,
Providence, Rhode Island, (1998): 1-36.


\noindent
Baez, John C., and James Dolan. "From Finite Sets to Feynman Diagrams"
Mathematics Unlimited - 2001 and Beyond, vol. 1, eds.
Björn Engquist and Wilfried Schmid, Springer, Berlin, (2001): 29-50.




%%------------------------
%
%% References Section
%\section*{REFERENCES}
%
%\noindent
%\newline
%\textbf{Reference Name} \\
%Full Title \\
%Address\\
%Email \\
%Telephone \\
%
%\noindent
%\textbf{Reference Name} \\
%Full Title \\
%Address\\
%Email \\
%Telephone \\
%
%\noindent
%\textbf{Reference Name} \\
%Full Title \\
%Address\\
%Email \\
%Telephone \\


% End document
\end{document}


